Predictive 3D modelling of free oblique cutting of Ti6Al4V titanium alloy and experimental validation for a wide range of conditions

### 2022-04-17 Article.tex
Changing the ANN figure for a PDF
Introduction of tables (instead of figures) and the new forces figures in the tex file
Text linked to the forces figures still to be updated

A 3-axis GF Mikron VCE 600 Pro milling machine is used to carry out dry free orthogonal and oblique cutting tests on Ti6Al4V (grade 5 annealed at 750°C for 1 h followed by air cooling) with the same kinematics as a shaper. As shown in Figure 1, the tungsten carbide tool (modified LCGN160602-0600-FG, CP500 from SECO) is fixed on a dedicated holder (modified CFHN-06 from SECO) and the sample to be cut is clamped in the spindle (no rotation is allowed during the test). The top of the sample includes 3 ribs of 1 mm in width (width of the tool is 6 mm) and 10 mm in length. The test consists of removing the upper layer (its height is the uncut chip thickness, h) of one rib at the prescribed cutting speed, $v_c$. The cutting speed is provided by the feed rate, $v_f$, of the machine (max. value of 40 m/min). The tool cutting edge inclination, $\lambda_s$, results from the orientation of the tool and the sample. Forces are measured with a 3-component Kistler 9257B dynamometer and are amplified by a Kistler 5070A charge amplifier. Acquisition is performed at 3 kHz with a Kistler 5697A2 data acquisition system and the DynoWare software. Table 1 shows the cutting conditions, each is repeated 3 times.

@@

\begin{figure}[h]
\centering
\caption{Tables 1 -- Cutting conditions of the study}
\includegraphics[width = 8cm]{Figures/Tab1}
\label{Tab1}
\end{figure}

@@

\begin{figure}[h]
\centering
\includegraphics{Figures/ANN}
\caption{Architecture of the ANN 3-9-7-1-sig used for the flow law}
\label{ANN}
\end{figure}

@@

FE modelling of the cutting process is very expensive in CPU computation time because of the coupling on many nonlinear phenomena and the large amount of tiny finite elements. Mass scaling (MS) is introduced in the model to reduce the CPU computation time while checking that it does not influence the results (forces and energies) via a mass scaling sensitivity study.
MS factors, ${MS}_f$, ranging from 1E6 (theoretical scaling of CPU time of 1000) to 1 (no scaling) have been used for one cutting condition ($\lambda_s$ = 0°, $v_c$ = 30 m/min and h = 60 µm). Table 2 gives the results normalised ($\hat{x}$) by these of the model without MS (x). As expected, actual speed-up does not increase linearly with the ${MS}_f$, but it is still significant. ${MS}_f$ of 1E6 leads to unstable computation and ${MS}_f$ of 1E5 results in erratic forces evolutions. These results are confirmed by high values of the kinetic ($KE$) on internal ($IE$) energies ratio (it should not exceed a few \%). A ${MS}_f$ value of 1E3 is selected as it provides a good balance between computation time reduction and impact on forces (considered in the analysis), while keeping $\frac{KE}{IE}$ below 1\%. To provide an order of magnitude of CPU computation time, between 10 h and 50 h are needed on 4 cores of an Intel i7-5700 HQ at 3.48 GHz.

\begin{figure}[h]
\centering
\caption{Tables 2 -- MS sensitivity study (selected ${MS}_f$ in bold)}
\includegraphics[width = 8cm]{Figures/Tab2}
\label{Tab2}
\end{figure}

@@

The number of neurons on the hidden layers may influence the results. A sensitivity study on the number of neurons for the first hidden layer is carried out to select the ANN providing the best balance between CPU computation time and quality of the results. The study shows no influence on the forces when compared to the built-in Johnson-Cook model, only computation time is influenced (Table 3). A first hidden layer with 9 neurons is therefore selected.

\begin{figure}[h]
\centering
\caption{Tables 3 -- Sensitivity to the number of neurons of the first layer, $\zeta$ (selection in bold)}
\includegraphics[width = 8cm]{Figures/Tab3}
\label{Tab3}
\end{figure}

@@

\begin{figure}[h]
\centering
\includegraphics{Figures/Fc}
\caption{Comparison of experimental (E) and numerical (N) cutting forces}
\label{Fc}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics{Figures/Ff}
\caption{Comparison of experimental (E) and numerical (N) feed forces}
\label{Ff}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics{Figures/Fp}
\caption{Comparison of experimental (E) and numerical (N) passive forces}
\label{Fp}
\end{figure}

>>>>

A 3-axis GF Mikron VCE 600 Pro milling machine is used to carry out dry free orthogonal and oblique cutting tests on Ti6Al4V (grade 5 annealed at 750°C for 1 h followed by air cooling) with the same kinematics as a shaper. As shown in Figure 1, the tungsten carbide tool (modified LCGN160602-0600-FG, CP500 from SECO) is fixed on a dedicated holder (modified CFHN-06 from SECO) and the sample to be cut is clamped in the spindle (no rotation is allowed during the test). The top of the sample includes 3 ribs of 1 mm in width (width of the tool is 6 mm) and 10 mm in length. The test consists of removing the upper layer (its height is the uncut chip thickness, h) of one rib at the prescribed cutting speed, $v_c$. The cutting speed is provided by the feed rate, $v_f$, of the machine (max. value of 40 m/min). The tool cutting edge inclination, $\lambda_s$, results from the orientation of the tool and the sample. Forces are measured with a 3-component Kistler 9257B dynamometer and are amplified by a Kistler 5070A charge amplifier. Acquisition is performed at 3 kHz with a Kistler 5697A2 data acquisition system and the DynoWare software. Table \ref{tab:CutCond} shows the cutting conditions, each is repeated 3 times.

@@

%
\begin{table}[!h]
\begin{center}
\caption{\label{tab:CutCond} Cutting conditions of the study}
\begin{tabular}{ll}
\hline\noalign{\smallskip}
Parameter  & Values\\
\noalign{\smallskip}\hline\noalign{\smallskip}
Cutting speed, $v_c$ (m/min) & 5, 7.5, 10, 20, 30, 40\\
Uncut chip thickness, $h$ (\textmu{}m) & 40, 60, 80\\
Cutting edge inclination, $\lambda_s$ & 0, 6\\
Width of the workpiece (mm) & 1\\
Length of the workpiece (mm) & 10\\
Width of the cutting edge (mm) & 6 (1.1 in the model)\\
Cutting edge radius, $r_\beta$ (\textmu{}m) & 20\\
Rake angle, $\gamma_0$ (\textdegree{}) & 15\\
Clearance angle, $\alpha_0$ (\textdegree{}) & 2\\
\noalign{\smallskip}\hline
\end{tabular}
\end{center}
\end{table}
%

@@

\begin{figure}[h]
\centering
\includegraphics[width=0.8\columnwidth]{Figures/ANN}
\caption{Architecture of the ANN 3-9-7-1-sig used for the flow law}
\label{ANN}
\end{figure}

@@

FE modelling of the cutting process is very expensive in CPU computation time because of the coupling on many nonlinear phenomena and the large amount of tiny finite elements. Mass scaling (MS) is introduced in the model to reduce the CPU computation time while checking that it does not influence the results (forces and energies) via a mass scaling sensitivity study.
MS factors, ${MS}_f$, ranging from 1E6 (theoretical scaling of CPU time of 1000) to 1 (no scaling) have been used for one cutting condition ($\lambda_s$ = 0°, $v_c$ = 30 m/min and h = 60 µm). Table \ref{tab:MS} gives the results normalised ($\hat{x}$) by these of the model without MS (x). As expected, actual speed-up does not increase linearly with the ${MS}_f$, but it is still significant. ${MS}_f$ of 1E6 leads to unstable computation and ${MS}_f$ of 1E5 results in erratic forces evolutions. These results are confirmed by high values of the kinetic ($KE$) on internal ($IE$) energies ratio (it should not exceed a few \%). A ${MS}_f$ value of 1E3 is selected as it provides a good balance between computation time reduction and impact on forces (considered in the analysis), while keeping $\frac{KE}{IE}$ below 1\%. To provide an order of magnitude of CPU computation time, between 10 h and 50 h are needed on 4 cores of an Intel i7-5700 HQ at 3.48 GHz.

%
\begin{table}[!h]
\begin{center}
\caption{\label{tab:MS} MS sensitivity study (selected $MS_f$ in bold)}
\begin{tabular}{lllll}
\hline\noalign{\smallskip}
$MS_f$  & Speed-up & $\hat{F_c}$ & $\hat{F_f}$ & $\frac{KE}{IE}$ (\%)\\
\noalign{\smallskip}\hline\noalign{\smallskip}
1 & 1 & 1 & 1 & 2.3E-4\\
1E2 & 9 & 1.006 & 0.982 & 2.2E-2\\
\textbf{1E3} & \textbf{21} & \textbf{1.008} & \textbf{0.940} & \textbf{2.2E-1}\\
1E4 & 61 & 1.012 & 0.921 & 2.4\\
1E5 & 173 & Erratic & Erratic & 22\\
1E6 & 207 & Unstable & Unstable & 58\\
\noalign{\smallskip}\hline
\end{tabular}
\end{center}
\end{table}
%

@@

The number of neurons on the hidden layers may influence the results. A sensitivity study on the number of neurons for the first hidden layer is carried out to select the ANN providing the best balance between CPU computation time and quality of the results. The study shows no influence on the forces when compared to the built-in Johnson-Cook model, only computation time is influenced (Table \ref{tab:NbNeurons}). A first hidden layer with 9 neurons is therefore selected.

%
\begin{table}[!h]
\begin{center}
\caption{\label{tab:NbNeurons} Sensitivity to the number of neurons of the first layer, $\zeta$
(selection in bold)}
\begin{tabular}{llll}
\hline\noalign{\smallskip}
$\zeta$  & Time increase (\%) & $\hat{F_c}$ & $\hat{F_f}$\\
\noalign{\smallskip}\hline\noalign{\smallskip}
Built-in & 0 & 1.000 & 1.000\\
\textbf{9} & \textbf{6} & \textbf{1.000} & \textbf{0.999}\\
11 & 6 & 1.001 & 1.000\\
13 & 7 & 1.000 & 0.998\\
15 & 8 & 1.001 & 1.001\\
17 & 10 & 1.000 & 1.000\\
\noalign{\smallskip}\hline
\end{tabular}
\end{center}
\end{table}
%

@@

\begin{figure}[h]
\centering
\includegraphics{Figures/Fx_0}
\caption{Comparison of experimental (E) and numerical (N) cutting forces at 0°}
\label{Fx_0}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics{Figures/Fx_6}
\caption{Comparison of experimental (E) and numerical (N) cutting forces at 6°}
\label{Fx_6}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics{Figures/Fy_0}
\caption{Comparison of experimental (E) and numerical (N) feed forces at 0°}
\label{Fy_0}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics{Figures/Fy_6}
\caption{Comparison of experimental (E) and numerical (N) feed forces at 6°}
\label{Fy_6}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics{Figures/Fz_6}
\caption{Comparison of experimental (E) and numerical (N) passive forces at 6°}
\label{Fz_6}
\end{figure}
